---
title: "Prediction and Time Series Computer Exercise Week 5"
author: 'Name: Nguyen Xuan Binh Student ID: 887799'
date: "12/02/2021"
output:
  word_document: default
  pdf_document: default
---

Exercise 5.5: The file t38.txt contains three quarterly time series. The time series start from the
first quarter of the year 1953 and the corresponding time series are,
CONS = total consumption (billions)
INC = income (billions)
INFLAT = inflation (%)
The time series CONS and INC represent the observed total consumption and income
in an imaginary country. The time series INFLAT represents inflation. The goal is
to estimate a so-called consumption function that explains the time series CONS with
the time series INC and INFLAT.
The conventional linear regression model for the response variable CONS is
CONS_t = β0 + β1 * INC_t + β2 * INFLAT_t + ε_t (5)
Assignments:
a) Estimate model (5) and study the goodness of fit.

```{r}
library(car)
library(forecast)
library(lmtest)
t38<-read.table("t38.txt",header=T,sep=" ")
cons<-t38$CONS
consTS <- ts(cons, start = 1953, frequency = 3)
inc<-t38$INC
incTS <- ts(inc, start = 1953, frequency = 3)
inflat<-t38$INFLAT
inflatTS <- ts(inflat, start = 1953, frequency = 3)
model5 <- lm (formula = cons ~ inc + inflat)
summary(model5)

```
Model study:
The regression coefficients corresponding to the income variable INC and inflation 
variable INFLAT are statistically significant with 5% level of significance.
The coefficient of the income is positive and the coefficient of the inflation is negative, indicating direction proportional relationship between consumption and inverse proportional relationship between consumption and inflation
The coefficient of determination is 0.9425, high enough to be considered as a good model

```{r}
par(mfrow = c(3,2),mar = c(3, 3, 3, 3))
# Q-Q plot of the residuals of model
qqnorm(model5$residuals,pch=16, main="Fig 1: Q-Q plot of the residuals of model 5")
qqline(model5$residuals,col="red",lwd=2)

#fitted model and original time series
fit <- ts(predict(model5), start = 1953, frequency = 3)
plot(consTS,col="red",main = "Fig 2: fitted model and original time series",xlab="Time",ylab="")
lines(fit,col="blue")
legend("topright", legend=c("cons", "fit"),
       col=c("red","blue"),lty=c(1,1),cex=0.5)
# Estimated residuals of model 5
plot(model5$residuals,type="p",main="Fig 3: Estimated residuals of model 5",ylab="Residuals",xlab="Year",pch=16,xaxt="n")
axis(1,at=seq(from=0,to=159,by=3),labels=seq(from=1953,to=2006,by=1))
abline(0,0)
# Cook’s distances of model 5
plot(cooks.distance(model5),main="Fig 4: Cook’s distances of model 5",ylab="Cook's distances",xlab="Year",pch=16,xaxt="n")
axis(1,at=seq(from=0,to=159,by=3),labels=seq(from=1953,to=2006,by=1))
# Estimated residuals of model 5
hist(model5$residuals,xlab="Residuals",ylab="Frequency",main="Fig 5: Estimated residuals of model 5")
# ACF of the estimated residuals of model 5
acf(model5$residuals, main = "Fig 6: ACF of the estimated residuals of model 5")
```

```{r}
# VIF of model 5
vif(model5)
```
Comments for the model 5:
• By Figures 5, the residuals seem to be normally distributed.
• By Figures 1 and 6, the residuals seem to be heavily correlated.
• By the variance inflation factor VIF, multicollinearity between CONS and INFLAT is unlikely as VIF is very close to 1.
• The reason for the correlatedness of the residuals can be seen from Figure 2,
where the fitted curve stays above and below the response variable CONS for
long time periods.
• There are many outliers in the cook's distance in figure 4
=> Model 5 can be considered to be inefficient in describing the model

b) Estimate the difference model corresponding to (5) and study the goodness of fit
```{r}
consD <- diff(cons)
incD <- diff(inc)
inflatD <- diff(inflat)
model5D<-lm(consD ~ incD + inflatD)
summary(model5D)
```
Model study:
The regression coefficients corresponding to the differenced income variable incD and differenced inflation variable inflatD are statistically significant with 5% level of significance.
The coefficient of the differenced income is positive and the coefficient of the differenced inflation is negative, still indicating their relationship with consumption found in (a) 
The coefficient of determination is 0.5826, suggesting that the differenced model may not be the correct model

```{r}
par(mfrow = c(3,2),mar = c(3, 3, 3, 3))
# Q-Q plot of the residuals of model
qqnorm(model5D$residuals,pch=16, main="Fig 1: Q-Q plot of the residuals of differenced model 5")
qqline(model5D$residuals,col="red",lwd=2)

#fitted model and original time series
fit <- ts(predict(model5D, start = 1953, frequency = 3))
plot(ts(consD),col="red",main = "Fig 2: differenced fitted model and differenced original time series",xlab="Time",ylab="")
lines(fit,col="blue")
legend("topright", legend=c("differenced cons", "differenced fit"),
       col=c("red","blue"),lty=c(1,1),cex=0.3)
# Estimated residuals of model 5
plot(model5D$residuals,type="p",main="Fig 3: Estimated residuals of differenced model 5",ylab="Residuals",xlab="Year",pch=16,xaxt="n")
axis(1,at=seq(from=0,to=159,by=3),labels=seq(from=1953,to=2006,by=1))
abline(0,0)
# Cook’s distances of model 5
plot(cooks.distance(model5D),main="Fig 4: Cook’s distances of differenced model 5",ylab="Cook's distances",xlab="Year",pch=16,xaxt="n")
axis(1,at=seq(from=0,to=159,by=3),labels=seq(from=1953,to=2006,by=1))
# Estimated residuals of model 5
hist(model5D$residuals,xlab="Residuals",ylab="Frequency",main="Fig 5: Estimated residuals of differenced model 5")
# ACF of the estimated residuals of model 5
acf(model5D$residuals, main = "Fig 6: ACF of estimated residuals of differenced model 5")
```
```{r}
vif(model5D)
```
Comments for the model 5 with differencing:
• By Figures 5, the residuals could be normally distributed.
• By Figures 1 and 6, the residuals are no longer heavily correlated.
• By the variance inflation factor VIF, multicollinearity between CONS diff and INFLAT diff is unlikely as VIF is very close to 1.
• There are many fewer outliers in the cook's distance in figure 4 compared to the undifferenced model
• In figure 2, the fit line does not match the differenced time series of consumption, suggesting that the differencing model is incorrect

```{r}
install.packages("lmtest")
library(lmtest)
model5_bg <- rep(NA,155)
# Breusch-Godfrey can be performed up to order:
# (sample size) - (number of estimated parameters) = 158-3 = 155
for (i in 1:155) {
  model5_bg[i]= bgtest(model5D, order=i)$p.value
}
which(model5_bg > 0.05)
# Null hypothesis of no autocorrelation accepted with all lags
```
- Judging by the Breusch-Godfrey test, there are many residuals with p_value > 0.05, proving to be correlated 
=> the null hypothesis of non-correlation is rejected
=> The differenced model 5 is inefficient to explain the relationship between response variable consumption and explanatory variables income and inflation

c) Estimate dynamic regression model:
CONS_t = β0 + β1 * CONS_t-1 + β2 * INC_t + β3 * INC_t-1 + β4 * INFLAT_t + β5 * INFLAT_t-1 + ε_t (6) and study the goodness of fit
```{r}
n <- nrow(t38)
model5dyna <- lm(consTS[-1] ~ consTS[-n] + incTS[-1] + incTS[-n] +
inflatTS[-1] + inflatTS[-n])
summary(model5dyna)
```
• All regression coefficients are statistically significant except inflat_t-1 variable

• The coefficient of the variable cons_t with lag 1 is 0.798, which implies that the
adjustment to changes in income and inflation is moderately fast.
• The signs of the coefficients of the income and inflation variables with lag
0 are as expected: the coefficient -0.79 of the inflation variable is negative and the
coefficient +0.49 of income variable is positive. These coefficients
describe the instant effects of changes in income and inflation.
• The signs of the coefficients of the income and inflation variables with lag
1 are also as expected.

• Interpretations of the regression coefficients of income and inflation variables with lag 0:
– If the income goes up by 1%, then total consumptions are instantly increased
by (without a lag) 0.498%.
– If the inflation is increased by 1%, then total consumptions are reduced by 0.793%.

• Interpretations of the long term elasticities of income and inflation variables:
– If the income goes up by 1%, then the total consumptions are reduced by 0.27%
in the long term.
– If the inflation is increased by 1%, then the total consumptions
are reduced by 0.25% in the long term.

• However, it is not possible to draw conclusions from the coefficient of the determination.

```{r}
par(mfrow = c(3,2),mar = c(3, 3, 3, 3))
# Q-Q plot of the residuals of model
qqnorm(model5dyna$residuals,pch=16, main="Fig 1: Q-Q plot of the residuals of dynamic regression model")
qqline(model5dyna$residuals,col="red",lwd=2)

#fitted model and original time series
fit <- ts(predict(model5dyna, start = 1953, frequency = 3))
plot(ts(cons),col="red",main = "Fig 2: dynamic fitted model and dynamic original time series",xlab="Time",ylab="")
lines(fit,col="blue")
legend("topright", legend=c("differenced cons", "differenced fit"),
       col=c("red","blue"),lty=c(1,1),cex=0.3)
# Estimated residuals of model 5
plot(model5dyna$residuals,type="p",main="Fig 3: Estimated residuals of dynamic regression model",ylab="Residuals",xlab="Year",pch=16,xaxt="n")
axis(1,at=seq(from=0,to=159,by=3),labels=seq(from=1953,to=2006,by=1))
abline(0,0)
# Cook’s distances of model 5
plot(cooks.distance(model5dyna),main="Fig 4: Cook’s distances of dynamic regression model",ylab="Cook's distances",xlab="Year",pch=16,xaxt="n")
axis(1,at=seq(from=0,to=159,by=3),labels=seq(from=1953,to=2006,by=1))
# Estimated residuals of model 5
hist(model5dyna$residuals,xlab="Residuals",ylab="Frequency",main="Fig 5: Estimated residuals of dynamic regression model")
# ACF of the estimated residuals of model 5
acf(model5dyna$residuals, main = "Fig 6: ACF of estimated residuals of dynamic regression model")
```
Comments related to the dynamic regression of model 5:
• By Figures 1 and 5, the residuals could be normally distributed.
• By Figures 3 and 6, the residuals are not correlated.
• By the residual diagrams in figure 3, there is no evidence of heteroscedasticity.
• By Figure 2, the fitted model coincides better with the original time series than
the fits of original model 5 found in (a) and differenced model 5 found in (b).

Now we test Breusch-Godfrey and VIF test
```{r}
library(lmtest)
model5dyna_bg <- rep(NA,152)
# Breusch-Godfrey can be performed up to order:
# (sample size) - (number of estimated parameters) = 158-6 = 152
for (i in 1:152) {
  model5dyna_bg[i]= bgtest(model5dyna, order=i)$p.value
}
which(model5dyna_bg < 0.05)
model5dyna_bg
vif(model5dyna)
```
• By the Breusch-Godfrey test, the residuals are not correlated. The null hypothesis
is accepted with 5% level of significance for all lags:
By VIF, there is strong multicollinearity in the model as all of the VIF values of the explanatory variables are bigger than 10 because the model involves same variables with different lags.

=> We consider this model to be sufficient in explaining the consumption variable

d) Which of the previous models are sufficient in explaining the behavior of the response
variable CONS?
The dynamic regression model found in (c) is an efficient model for the relationship between the response variable consumption and the explanatory variables income and inflation, as proven in (c)

